
---
title: "Global Life Expectancy Analysis"
---

This project explores global trends in life expectancy using World Bank data.

## Load Libraries
```{r}
library(tidyverse)
```


---

### 📥 Load Dataset
```{r}
df_raw <- read_csv("~/health-data-portfolio/projects/project1/API_SP.DYN.LE00.IN_DS2_en_csv_v2_121663.csv", skip = 4)
df_raw
```



---

## Cleaning Data
```{r}
df_clean <- df_raw %>%
  select(`Country Name`, starts_with("20")) %>%
  pivot_longer(
    cols = -`Country Name`,
    names_to = "year",
    values_to = "life_expectancy"
  ) %>%
  mutate(year = as.numeric(year)) %>%
  drop_na()
df_clean
```


---

##Global Trend Visualization
```{r}
df_global <- df_clean %>%
  group_by(year) %>%
  summarise(mean_life = mean(life_expectancy, na.rm = TRUE))

ggplot(df_global, aes(x = year, y = mean_life)) +
  geom_line(color = "blue", size = 1) +
  labs(
    title = "Global Life Expectancy Trend",
    x = "Year",
    y = "Life Expectancy"
  )
```



---

## Insight

Global life expectancy has shown a consistent increasing trend over the last two decades.

This improvement may reflect better access to healthcare, vaccination programs, and public health interventions across countries.

---

## Indonesia vs Australia vs Brazil
```{r}
selected_countries <- df_clean %>%
  filter(`Country Name` %in% c("Indonesia", "Australia", "Brazil"))

ggplot(selected_countries, aes(year, life_expectancy, color = `Country Name`)) +
  geom_line(size = 1) +
  labs(
    title = "Life Expectancy Comparison",
    y = "Life Expectancy"
  )
```








